\bf Are you normal? A new projection pursuit index to assess a sample against a multivariate null distribution

### 2024-06-27 paper.qmd
integrated Annalisa change

The projection of an ellipsoid onto \gD{} is an ellipse, where the curve of the ellipse is defined through the set of points $\mathbfit{x}$ for which the gradient is parallel to the projection plane. We call points in the projection that are on the curve $\mathbfit{y}$. From this we can compute the analog equation for the projection as

$$(\mathbfit{y} - \mathbfit{\mu}_p)(P^T \Sigma P)^{-1}(\mathbfit{y} - \mathbfit{\mu}_p)^T = c^2$$
where $P$ is a $(p\times 2)$ orthonormal basis defining the projection and $\mathbfit{\mu}_p = \mathbfit{\mu} P$ the projected mean. This means the matrix $(P^T \Sigma P)^{-1}$ is defining the ellipse in the \gD{} projection. In general $c$ could be any constant, but typically we would select it as a quantile of the $\chi^2$ distribution, so that the size of the ellipse corresponds to a selected probability.

>>>>

The projection of an ellipsoid onto \gD{} is an ellipse, where the curve of the ellipse is defined through the set of points $\mathbfit{x}$ for which the gradient is parallel to the projection plane. That is, the curve 
consists of $\mathbfit{x}$ satisfying

$$(\mathbfit{x}-\mathbfit{\mu}) \Sigma^{-1}(\mathbfit{x}-\mathbfit{\mu})^T = 2 (\mathbfit{x}-\mathbfit{\mu}) \Sigma^{-1} = 2\mathbfit{s}P^T
$$

for some $\mathbfit{s} \in \mathbb{R}^2$, where $P$ is a $(p\times 2)$ orthonormal basis defining the
projection. We can write $\mathbfit{x}-\mathbfit{\mu} = \mathbfit{s} P^T \Sigma$.
Making this substitution in the \pD{} ellipsoid equation yields

$$c^2 = (\mathbfit{x}-\mathbfit{\mu}) \Sigma^{-1}(\mathbfit{x}-\mathbfit{\mu})^T = \mathbfit{s} P^T \Sigma P \mathbfit{s}^T
$$

We call points in the projection that are on the curve $\mathbfit{y}$. From this we can compute the analog equation for the projection as

$$(\mathbfit{y} - \mathbfit{\mu}_p)(P^T \Sigma P)^{-1}(\mathbfit{y} - \mathbfit{\mu}_p)^T = c^2.$$
This means the matrix $(P^T \Sigma P)^{-1}$ is defining the ellipse in the \gD{} projection. In general $c$ could be any constant, but typically we would select it as a quantile of the $\chi^2$ distribution, so that the size of the ellipse corresponds to a selected probability.

### 2024-06-27 paper.tex
integrated Annalisa change

The projection of an ellipsoid onto \gD{} is an ellipse, where the curve
of the ellipse is defined through the set of points \(\mathbfit{x}\) for
which the gradient is parallel to the projection plane. We call points
in the projection that are on the curve \(\mathbfit{y}\). From this we
can compute the analog equation for the projection as

\[(\mathbfit{y} - \mathbfit{\mu}_p)(P^T \Sigma P)^{-1}(\mathbfit{y} - \mathbfit{\mu}_p)^T = c^2\]
where \(P\) is a \((p\times 2)\) orthonormal basis defining the
projection and \(\mathbfit{\mu}_p = \mathbfit{\mu} P\) the projected
mean. This means the matrix \((P^T \Sigma P)^{-1}\) is defining the
ellipse in the \gD{} projection. In general \(c\) could be any constant,
but typically we would select it as a quantile of the \(\chi^2\)
distribution, so that the size of the ellipse corresponds to a selected
probability.

>>>>

The projection of an ellipsoid onto \gD{} is an ellipse, where the curve
of the ellipse is defined through the set of points \(\mathbfit{x}\) for
which the gradient is parallel to the projection plane. That is, the
curve consists of \(\mathbfit{x}\) satisfying

\[(\mathbfit{x}-\mathbfit{\mu}) \Sigma^{-1}(\mathbfit{x}-\mathbfit{\mu})^T = 2 (\mathbfit{x}-\mathbfit{\mu}) \Sigma^{-1} = 2\mathbfit{s}P^T
\]

for some \(\mathbfit{s} \in \mathbb{R}^2\), where \(P\) is a
\((p\times 2)\) orthonormal basis defining the projection. We can write
\(\mathbfit{x}-\mathbfit{\mu} = \mathbfit{s} P^T \Sigma\). Making this
substitution in the \pD{} ellipsoid equation yields

\[c^2 = (\mathbfit{x}-\mathbfit{\mu}) \Sigma^{-1}(\mathbfit{x}-\mathbfit{\mu})^T = \mathbfit{s} P^T \Sigma P \mathbfit{s}^T
\]

We call points in the projection that are on the curve \(\mathbfit{y}\).
From this we can compute the analog equation for the projection as

\[(\mathbfit{y} - \mathbfit{\mu}_p)(P^T \Sigma P)^{-1}(\mathbfit{y} - \mathbfit{\mu}_p)^T = c^2.\]
This means the matrix \((P^T \Sigma P)^{-1}\) is defining the ellipse in
the \gD{} projection. In general \(c\) could be any constant, but
typically we would select it as a quantile of the \(\chi^2\)
distribution, so that the size of the ellipse corresponds to a selected
probability.
